---
title: "Stats 506, F20, Final Project"
author: "Zhilin He, zhilinhe@umich.edu"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
# 79: -------------------------------------------------------------------------
#! We generally don't need to see the code in the results document.
knitr::opts_chunk$set(echo = FALSE)
#! Make sure all chunks have a label. This one is labelled 'setup'.
```

# Link to GitHub repository
https://github.com/zhilinheobv/506-Final-Project

# Introduction

Fuel oil (including kerosene and diesel) is a common source of energy. Although
it is not as environmental friendly as natural gas, it is used for heating in
many homes and buildings, and other industries like fueling for trucks and 
ships. Some industries aim to reduce the use of polluting fuel oil since 
it might cause air pollution and harm people's health<sup>[1]</sup>. Therefore,
proper usage of fuel oil is an interesting topic to consider.

Many buildings use fuel oil as an energy source for various reasons. It's
possible that buildings' ownership affects the choice of energy sources. In
this project, we are trying to find out whether government-owned commercial
buildings use more fuel oil than not government-owned buildings, and what fuel
oil is used for in different buildings.

We are going to use veral different statistical methods including Pearson's
chi-squared test, two sample t-test, propensity score analysis to 
identify the difference between goverment-owned buildings and not 
government-owned ones. 

# Data

This project uses the 2012 CBECS data. It contains 1119 variables and 6720 
sampled observations. The data contains all aspects of information concerning
energy consumption about commercial buildings all across the US.

We choose Detailed descriptions for the chosen variables are shown in Table 1 below.

| Variable Name   | Variable Description                           |
| --------------- |:----------------------------------------------:|
| PUBID (id)      | Building ID                                    |
| REGION          | Census region                                  |
| CENDIV (div)    | Census division                                |
| GOVOWN (gov)    | Government owned (1) or not (2)                |
| FKUSED          | Uses fuel oil/diesel/kerosene (1) or not (2)   |
| FKTYPE          | Fuel oil, diesel or kerosene used              |
| FKHTBTU         | Fuel oil^ used for heating (thousand BTU)      |
| FKCLBTU         | Fuel oil used for cooling (thousand BTU)       |
| FKWTBTU         | Fuel oil used for water heating (thousand BTU) |
| FKCKBTU         | Fuel oil used for cooking (thousand BTU)       |
| FKOTBTU         | Fuel oil used for other reasons (thousand BTU) |
| FKBTU           | Annual fuel oil consumption (thousand BTU)     |
| FKEXP           | Annual fuel oil expenditure ($)                |
| FINALWT         | Full sample building weight                    |
| FINALWT1-197    | Final replicate weights                        |

^ Including diesel and kerosene. Same with "Fuel oil" below.

<center>**Table 1**. Variable descriptions</center>

# Methods

We use the jackknife method to estimate standard error. This method uses
replicate weights to repeatedly estimate the statistics of interest and 
calculate the differences between these estimates and the full-sample 
estimate<sup>[2]</sup>.

We also use Pearson's chi-squared test to verify whether the effect of
building ownership is significant for fuel oil usage.

# Analysis

## Data Cleaning

Firstly, we take a look at the data. Data summary is shown below.

```{r load_data, include=FALSE}
library(survey)
library(ggplot2)
library(tidyverse)
library(knitr)
cbecs = read.csv("./2012_public_use_data_aug2016.csv")
```

```{r data cleaning, echo=FALSE, class.output = ".bordered"}
## Select variables and clean the data
cbsdt = cbecs %>%
  mutate(gov=factor(GOVOWN), div=factor(CENDIV), fo_used=factor(FKUSED)) %>%
  select(id=PUBID, div, gov, fo_used, fo_type=FKTYPE, heating=FKHTBTU, 
         cooling=FKCLBTU, waterheating=FKWTBTU, cooking=FKCKBTU, 
         other=FKOTBTU, total=FKBTU, expenditure=FKEXP) %>%
  mutate(fo_type = factor(ifelse(fo_used == 1, 
                                 ifelse(fo_type %in% 4:7, 4, fo_type), NA)))
# There is 1 wrong observation for the FKTYPE variable.
cbswt = cbecs %>% select(starts_with("FINALWT")) # Replicate weights
levels(cbsdt$div)=c('New England', 'Middle Atlantic', 'East North Central',
                    'West North Central', 'South Atlantic', 
                    'East South Central', 'West South Central',
                    'Mountain', 'Pacific')
levels(cbsdt$gov) = c('Goverment', 'Non-goverment')
levels(cbsdt$fo_used) = c(TRUE, FALSE)
levels(cbsdt$fo_type) = c('Fuel oil', 'Diesel', 'Kerosene', 'More than one',
                          'Unknown')
summ = data.frame(unclass(summary(cbsdt)), check.names = FALSE, 
                  stringsAsFactors = FALSE)
summ[is.na(summ)] = ''
kable(summ[, 1:6], row.names=F)
kable(summ[, 7:12], row.names=F)
```

<center>**Table 2**. Summary of variables</center>

## Fuel oil usage

We calculate the weighted frequency of fuel oil usage using the survey
weights.

```{r usage, echo=FALSE}
sampweights = cbswt$FINALWT
wts = cbswt[, -1]
des = svrepdesign(weights=sampweights, repweights=wts, type="JK1", 
                  scale = 4/197, rscales = rep(1, 197), mse=TRUE,
                  data=cbsdt)
psumm = svyby(~fo_used, by=~gov+div, des, svytotal) %>%
  mutate(percent = fo_usedTRUE / (fo_usedTRUE + fo_usedFALSE))
p1 = ggplot(psumm, aes(x=div, y=percent, fill=gov)) +
      geom_bar(position="dodge", stat="identity") +
      scale_x_discrete(guide=guide_axis(n.dodge=2)) +
      xlab('Division') +
      ylab('Percent of buildings using fuel oil') +
      theme(legend.title = element_blank())
ptotal = svyby(~fo_used, by=~gov, des, svytotal)
ps = psumm %>% rename(ownership=gov, division=div, `percent used`=percent,
                      "fuel oil used" = fo_usedTRUE,
                      "fuel oil not used" = fo_usedFALSE) %>%
               mutate(`LB used` = `fuel oil used`-qnorm(.975)*se1,
                      `UB used` = `fuel oil used`+qnorm(.975)*se1,
                      `LB not used` = `fuel oil not used`+qnorm(.975)*se2,
                      `UB not used` = `fuel oil not used`+qnorm(.975)*se2) %>%
               select(-c(se1, se2))
kable(ps, row.names=F)
```

<center>**Table 3**. Frequency table of buildings using oil
by census division and ownership</center>

```{r fig1, echo=FALSE}
p1
```

<center>**Figure 1**. Percent of buildings using fuel oil by 
census division and ownership</center>

From figure 1, we can see that the percentage of government buildings that
use fuel oil is greater in all divisions, but the difference is more obvious
in Middle Atlantic, East North Central, West North Central and South Atlantic
divisions. We can use Pearson chi-squared test to verify whether this
difference is statistically significant.

```{r chisq_test, echo=FALSE}
pchi = data.frame(matrix(ncol = 3, nrow = 10))
names(pchi) = c("Division", "Chi-squared statistic", "p-value")
pchi$Division = c(levels(cbsdt$div), "Total")
a = chisq.test(ptotal[, 2:3])
for(i in 1:9) {
  res = chisq.test(psumm[(2 * i - 1):(2 * i), 3:4])
  pchi[i, 2] = res$statistic
  pchi[i, 3] = ifelse(res$p.value < 0.001, "p < 0.001", round(res$p.value, 3))
}
pchi[10, 2] = a$statistic
pchi[10, 3] = ifelse(a$p.value < 0.001, "p < 0.001", round(a$p.value, 3))
kable(pchi, row.names=F)
```

<center>**Table 4**. Result of chi-squared tests</center>

It turns out the difference of fuel oil usage between government and 
non-government buildings is significant for all divisions. This is
possibly due to the fact that the full sample size is large enough to
identify this difference.

## Consumption

We calculate the annual consumption of fuel oil among the buildings using
fuel oil as an energy source. The consumptions for each specific use are
also calculated. The graph for consumptions in different divisions is
shown below, with estimated error bars:

```{r consumption, echo=FALSE}
fouseddf = cbind(cbswt, cbsdt) %>% filter(fo_used == TRUE)
sampweights1 = fouseddf$FINALWT
wts1 = fouseddf[, 2:198]
des1 = svrepdesign(weights=sampweights1, repweights=wts1, type="JK1", 
                  scale = 4/197, rscales = rep(1, 197), mse=TRUE,
                  data=fouseddf)
pconsm = svyby(~total, by=~gov+div, des1, svymean) %>%
  mutate(lower = total - se * qnorm(.975),
         upper = total + se * qnorm(.975)) %>%
  rename(mean = total)
p2 = ggplot(pconsm, aes(x=div, y=mean, fill=gov)) +
      geom_bar(position="dodge", stat="identity") +
      geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                    position=position_dodge(.9)) +
      scale_x_discrete(guide=guide_axis(n.dodge=2)) +
      xlab('Division') +
      ylab('Mean annual consumption of fuel oil (thousand BTU)') +
      theme(legend.title = element_blank())
p2
```

<center>**Figure 2**. Mean annual fuel oil consumption for buildings
by division and ownership</center>

```{r consumption_total, echo=FALSE}
ptotcon = svyby(~total, by=~gov, des1, svymean) %>%
  mutate(lower = total - se * qnorm(.975),
         upper = total + se * qnorm(.975)) %>%
  rename(mean = total)
ptotcon$div = "Total"
pcon = rbind(pconsm, ptotcon)
kable(pcon, row.names=F)
```

<center>**Table 5**. Mean annual fuel oil consumption for buildings</center>

In New England, Middle Atlantic, East North Central and South Atlantic 
divisions, government buildings consume significantly more fuel oil than
non-government buildings on average. In Mountain Division, the difference is
not significant. In other divisions, non-government buildings consume more fuel
oil annually. Overall, because New England and Middle Atlantic divisions 
consume fuel oil the most in the US, government buildings consume significantly
more fuel oil than other buildings annually on average.

## Detailed uses

Fuel oil can be used in many ways in buildings. We consider the differences of
uses between goverment buildings and non-government ones.

```{r uses, echo=FALSE}
puses = svyby(~heating + cooling + waterheating + cooking + other,
              by=~gov+div, des1, svymean) 
ptuses = svyby(~heating + cooling + waterheating + cooking + other,
              by=~gov, des1, svymean) 
ptuses$div = "Total"
pus = rbind(puses, ptuses) %>% select(-starts_with("se"))
kable(pus, row.names=F)
```

<center>**Table 6**. Detailed uses of fuel oil</center>

```{r fig3, echo=FALSE, fig.width = 9, fig.height = 15}
p3 = pus %>% 
  pivot_longer(-c(gov, div), names_to = "use", values_to = "consumption") %>%
  ggplot(aes(fill=use, x=gov, y=consumption)) +
   geom_bar(position="stack", stat="identity") +
   ylab('Annual onsumption (thousand BTU)') +
   xlab('Division') +
   facet_wrap(~div)
p3
```

<center>**Figure 3**. Stacked bar plots for different uses of fuel oil</center>

From the figure we can clearly see that uses of fuel oil mainly differs by
census divisions, possibly due to climate reasons. The uses of fuel oil in
government-owned and not government-owned buildings are not so different
if heating is the main use of fuel oil in the division. For East South Central,
West South Central, Mountain and Pacific divisions, heating is not the main 
use of fuel oil. In these divisions, government buildings almost only use fuel
oil for other reasons (including manufacturing and electricity generation).
But non-government buildings still use some fuel oil for heating.

## Type of fuel oil

In this project, "fuel oil" includes (heavy) fuel oil, kerosene and diesel.
We consider the differences of type of fuel oil used in those buildings
using fuel oil as an energy source.

```{r type_fuel_oil, echo=FALSE}
ptype = svyby(~fo_type, by=~gov+div, des1, svytotal) 
pttype = svyby(~fo_type, by=~gov, des1, svytotal)
pttype$div = "Total"
ptp = rbind(ptype, pttype) %>% select(-starts_with("se"))
kable(ptp, row.names=F)
```

<center>**Table 7**. Types of fuel oil</center>

```{r fig4, echo=FALSE, fig.width = 9, fig.height = 15}
p4 = ptp %>%
  pivot_longer(-c(gov, div), names_prefix = "fo_type",
               names_to = "type", values_to = "count") %>%
  ggplot(aes(fill=type, x=gov, y=count)) +
   geom_bar(position="fill", stat="identity") +
   ylab('Percent') +
   xlab('Division') +
   facet_wrap(~div)
p4
```

<center>**Figure 4**. Stacked bar plots for type of fuel oil </center>

In general, government buildings use less heavy fuel oil than non-government
buildings. However, in Mountain Division, government buildings use 
significantly more heavy fuel oil. In New England, East and West North Central
divisions, government buildings use kerosene more, but in other divisions,
the difference is not obvious or non-government buildings use kerosene more.
The type of fuel oil that buildings use also differs by census division.

# Conclusion

The analysis about CBECS 2012 data provides some evidence for the assumption 
that government-owned buildings are more likely to use fuel oil as a source of
energy, and they consume more fuel oil. However, generally non-government
buildings consume more heavy fuel oil which is more polluting than kerosene and
diesel. 

Different census divisions use fuel oil for different reasons. In divisions
where heating is not the main use of fuel oil, government buildings don't use
fuel oil for heating entirely, while non-government buildings still use it for
heating.

# To-do list

1. Finish conclusion.

2. Possibly add analysis by considering the effect of building activity.

3. Possibly add propensity score analysis.

# References

[1] [Fuel oil](en.wikipedia.org/wiki/Fuel_oil). Wikipedia.

[2] [Jackknife resampling](https://en.wikipedia.org/wiki/Jackknife_resampling).
Wikipedia.

[3] [Propensity Score Analysis with Survey Weighted Data](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5802372/). 